Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone

### 2021-05-06 paper/ifo-appendix.tex
minor changes from read through
comments from Andrew
adding Changrong's code, updating README, misc comments from Andrew

\section{Open-source code}
\label{app:code}
This project is implemented in Python 3~\cite{python} scripts and jupyter notebooks~\cite{jupyter,ipython} as well as in MATLAB~\cite{MATLAB}. We used the Python packages of numpy~\cite{numpy}, scipy~\cite{scipy}, matplotlib~\cite{matplotlib}, tqdm~\cite{tqdm}, and logmmse~\cite{logmmse}.
The current build and sample data can be found at:
\textbf{[Removed GitHub web-link for double-anonymous review.]}
%\url{https://github.com/daccordeon/gravexplain}
% \begin{verbatim}
% README
% \end{verbatim}

@@

\section{Detecting a sinusoidal signal in Gaussian noise}
\label{app:sinusoid_likelihood}
%\jam{[to do! from Andrew: ``Write a short appendix proving that maximizing the likelihood exp(-n||n/2) with respect to A given a signal A cos (wt + phi) is the Fourier transform. Classic sig proc result useful for teachers to know.'']}
In this appendix, we demonstrate the modulus of the Fourier transform is an appropriate detection statistic when searching for a sinusoidal signal in Gaussian noise. 
We describe the data as
\begin{equation}
x(t) = s(t) + n(t)\,, 
\end{equation}
where $s(t)$ and $n(t)$ are the signal and noise respectively. 
Then we describe the signal as 
\begin{equation}
s(t) = A \cos{\omega t + \phi}\,,
\end{equation}
where $A$, $\omega$, and $\phi$ are the amplitude, angular frequency and phase of the signal respectively. 
The log-likelihood assuming Gaussian noise is given by
\begin{eqnarray}
\log \mathcal{L} &=& -\frac{1}{2}\ip{n}{n}\,, \\
                 &=& -\frac{1}{2}\ip{x-s}{x-s}\,, \\ 
                 &=& -\frac{1}{2}\ip{x}{x} - \frac{A^2}{2} \ip{\cos[\omega t + \phi]}{\cos[\omega t + \phi]} \nonumber\\
                 &&+ A \ip{x}{\cos[\omega t + \phi]}\,,\\ 
                 &=& -\frac{1}{2}\ip{x}{x} - \frac{A^2}{4} + A \ip{x}{\cos[\omega t + \phi]}\,,\label{eqn:logLike}
\end{eqnarray}
where we define the inner product between $u(t)$ and $v(t)$ as $\ip{u}{v} = \int_0^T \mathrm{d} t \, u(t) v(t) $ and $T$ is the total time of the data (and the last line assumes that $T\gg1/\omega$ so that $\ip{\cos[\omega t + \phi]}{\cos[\omega t + \phi]}\approx{1/2}$).

Maximising Eqn.~\ref{eqn:logLike} with respect to $A$ shows that the maximum likelihood is when $A = 2\ip{x}{\cos[\omega t + \phi]}$. 
The maximum likelihood is then, 
\begin{eqnarray}
\label{eqn:maxLogLike}
\log \mathcal{L}_\mathrm{max} = - \frac{1}{2} \ip{x}{x} + \ip{x}{\cos[\omega t + \phi]}^2\,. 
\end{eqnarray}
Focusing on the second term of Eqn~\ref{eqn:maxLogLike}, we see that the maximum log-likelihood of a sinusoidal signal in Gaussian noise is the modulus of the cosine Fourier transform, baring the constant factor $\ip{x}{x}$.

@@

%\subsection{Pseudocode}
Here we give some pseudocode of the implementation used in Section~\ref{sec:viterbi}. More general pseudocode complete with technical language is available elsewhere online~\cite{viterbipseudocode:online}. We use Fourier amplitudes, normalized between $(0, 1)$ by dividing by the maximum value in the grid, as multiplicative weights, but to avoid numerical underflow we take the logarithm and consider them as additive weights.

Let $X$ be grid of $j=0,\ldots,N_f$ rows and $i=0,\ldots,N_t$ columns of additive weights for each node. Let Y and Z be grids of the same shape to store the weight of the best path to each node and the row index of the previous node on that path, respectively. Let W be a length $N_t+1$ array to store the final sequence of row indices for the optimal path. We restrict paths to only move one cell up or down at a time (or stay constant).

@@

	    	\State $Y_{i,j} \gets \underset{k \in \{-1,0,1\}}{\max} (Y_{i-1,j+k} + X_{i,j})$
	    	\State $Z_{i,j} \gets j + \underset{k \in \{-1,0,1\}}{\arg\max} (Y_{i-1,j+k} + X_{i,j})$

>>>>

\section{Open-source code}
\label{app:code}
This project is implemented in Python 3~\cite{python} scripts and jupyter notebooks~\cite{jupyter,ipython} as well as in MATLAB~\cite{MATLAB}. We used the Python packages of numpy~\cite{numpy}, scipy~\cite{scipy}, matplotlib~\cite{matplotlib}, tqdm~\cite{tqdm}, and logmmse~\cite{logmmse}.
The current build and sample data can be found at:
%\textbf{[Removed GitHub web-link for double-anonymous review.]}
\url{https://github.com/daccordeon/gravexplain}
% \begin{verbatim}
% README
% \end{verbatim}

@@

\section{Detecting a sinusoidal signal in Gaussian noise}
\label{app:sinusoid_likelihood}
%\jam{[to do! from Andrew: ``Write a short appendix proving that maximizing the likelihood exp(-n||n/2) with respect to A given a signal A cos (wt + phi) is the Fourier transform. Classic sig proc result useful for teachers to know.'']}
In this appendix, we demonstrate the modulus of the Fourier transform is an appropriate detection statistic when searching for a sinusoidal signal in Gaussian noise. 
We describe the data as
\begin{equation}
x(t) = s(t) + n(t)\,, 
\end{equation}
where $s(t)$ and $n(t)$ are the signal and noise respectively. 
Then we describe the signal as 
\begin{equation}
s(t) = A \cos\left[{\omega t + \phi}\right]\,,
\end{equation}
where $A$, $\omega$, and $\phi$ are the amplitude, angular frequency and phase of the signal respectively. 
The log-likelihood assuming Gaussian noise is given by
\begin{eqnarray}
\log \mathcal{L} &=& -\frac{1}{2}\ip{n}{n}\,, \\
                 &=& -\frac{1}{2}\ip{x-s}{x-s}\,, \\ 
                 &=& -\frac{1}{2}\ip{x}{x} - \frac{A^2}{2} \ip{\cos[\omega t + \phi]}{\cos[\omega t + \phi]} \nonumber\\
                 &&+ A \ip{x}{\cos[\omega t + \phi]}\,,\\ 
                 &=& -\frac{1}{2}\ip{x}{x} - \frac{A^2}{4} + A \ip{x}{\cos[\omega t + \phi]}\,,\label{eqn:logLike}
\end{eqnarray}
where we define the inner product between $u(t)$ and $v(t)$ as $\ip{u}{v} = \int_0^T \mathrm{d} t \, u(t) v(t) $ and $T$ is the total time of the data (and the last line assumes that $T\gg1/\omega$ so that $\ip{\cos[\omega t + \phi]}{\cos[\omega t + \phi]}\approx{1/2}$).

Maximising Eqn.~\ref{eqn:logLike} with respect to $A$ shows that the maximum likelihood is when $A = 2\ip{x}{\cos[\omega t + \phi]}$. 
The maximum likelihood is then, 
\begin{eqnarray}
\label{eqn:maxLogLike}
\log \mathcal{L}_\mathrm{max} = - \frac{1}{2} \ip{x}{x} + \ip{x}{\cos[\omega t + \phi]}^2\,. 
\end{eqnarray}
From the second term of Eqn~\ref{eqn:maxLogLike}, we see that the maximum log-likelihood of a sinusoidal signal in Gaussian noise is the modulus of the cosine Fourier transform, plus the term $\ip{x}{x}$, which is independent of $\omega$ and $\phi$ and can therefore be ignored when searching over $\omega$.
%baring the constant factor $\ip{x}{x}$.

@@

%\subsection{Pseudocode}
Here we present some pseudocode of the implementation used in Section~\ref{sec:viterbi}. More general pseudocode complete with technical language is available elsewhere online~\cite{viterbipseudocode:online}. We use Fourier amplitudes, normalized between $(0, 1)$ by dividing by the maximum value in the grid, as multiplicative weights. To avoid numerical underflow we take the logarithm and consider them instead as additive weights.

Let $X$ be grid of $j=0,\ldots,N_f$ rows and $i=0,\ldots,N_t$ columns of additive weights for each node. Let Y and Z be grids of the same shape to store the weight of the best path to each node and the row index of the previous node on that path, respectively. Let W be a length $N_t+1$ array to store the final sequence of row indices for the optimal path. We restrict paths to only move one cell up or down at a time (or stay constant). For the edge cases of $j=0,N_f$, we only search over $k \in \{0,1\}$ and $k \in \{-1,0\}$, respectively, to stay inside the grid (this is not shown in the pseudocode).

@@

	    	\State $Y_{i,j} \gets X_{i,j} + \underset{k \in \{-1,0,1\}}{\max} (Y_{i-1,j+k})$
	    	\State $Z_{i,j} \gets j + \underset{k \in \{-1,0,1\}}{\arg\max} (Y_{i-1,j+k})$

### 2021-05-06 paper/ifo-complexAudio.tex
adding Changrong's code, updating README, misc comments from Andrew

In signal processing, the ideal filter would be one that:
(i) completely attenuates the undesired parts of the spectrum, 
(ii) does not change the rest of the spectrum, and 
(iii) smoothly transitions between these regions, as to not damage the time domain signal when seen under convolution. 
However, these three conditions cannot all hold at once. For example, if conditions (i) and (ii) hold, then the filter must be discontinuous at the boundary of the undesired region but this implies that the filter has ``infinite latency'' and so will affect (or damage) the time domain signal for infinite time~\cite{10.5555/151045}. Therefore, any filter must compromise between these three conditions. For speech intelligibility, this means that either: (i) some noise remains in the filtered recording, (ii) some of the speech content is lost as certain important frequencies are attenuated, or (iii) the speech is somewhat distorted in time. All three of these cases can, in extreme cases, worsen speech intelligibility beyond the unfiltered recording and so we choose filters that compromise between achieving the three conditions.
% https://en.wikipedia.org/wiki/Filter_(signal_processing)

>>>>

% https://en.wikipedia.org/wiki/Filter_(signal_processing)
In signal processing, the ideal filter would be one that:
(i) completely attenuates the undesired parts of the spectrum, 
(ii) does not change the rest of the spectrum, and 
(iii) smoothly transitions between these regions, as to not damage the time domain signal when seen under convolution. 
However, these three conditions cannot all hold at once. For example, if conditions (i) and (ii) hold, then the filter must be discontinuous at the boundary of the undesired region but this implies that the filter has ``infinite latency'' and so will affect (or damage) the time domain signal for infinite time~\cite{10.5555/151045}. Therefore, any filter must compromise between these three conditions. For speech intelligibility, this means that either: (i) some noise remains in the filtered recording, (ii) some of the speech content is lost as certain important frequencies are attenuated, or (iii) the speech is somewhat distorted in time.
All three of these cases can, when taken to the extreme, make the speech intelligibility worse than the unfiltered recording, i.e.\ the results can be worse than doing no filtering. Therefore, we choose filters that compromise between achieving the three conditions.
% All three of these cases can, in extreme cases, worsen speech intelligibility beyond the unfiltered recording and so we choose filters that compromise between achieving the three conditions.
